General Game Playing and Monte-Carlo Tree Search

### 2013-11-15 doc/paper/paper.tex
code blocks for GDL and core.logic

\documentclass[letterpaper]{article}
% Required Packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\nocopyright
% Title
\title{General Game Playing and Monte-Carlo Tree Search}
\author{Armando Ramirez \\ Purdue University, West Lafayette, IN \\ ramirez7@purdue.edu \\ armando.m.ramirez@gmail.com}
% Body
\begin{document}
\maketitle

@@

\subsubsection{core.logic}
%FIXME CAN'T LINEBREAK AFTER THIS SUBSUBSECTION
The primary Clojure library used in the implementation was core.logic, which is a port of the logic programming language miniKanren as formulated by William Byrd in his 2010 PhD thesis \cite{byrd2010relational}.
\subsection{GDL Transformation to core.logic}

>>>>

\documentclass[letterpaper]{article}
% Required Packages
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{listings}
\lstset{language=Lisp, basicstyle=\footnotesize}
\usepackage{caption}
\DeclareCaptionFont{white}{ \color{white} }
\captionsetup[lstlisting]{singlelinecheck=false, margin=0pt, font={bf,footnotesize} }
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\nocopyright
% Title
\title{General Game Playing and Monte-Carlo Tree Search}
\author{Armando Ramirez \\ Purdue University, West Lafayette, IN \\ ramirez7@purdue.edu \\ armando.m.ramirez@gmail.com}
% Body
\begin{document}
\maketitle

@@

\subsubsection{core.logic}
%FIXME CAN'T LINEBREAK AFTER THIS SUBSUBSECTION
The primary Clojure library used in the implementation was core.logic, which is a port of the logic programming language miniKanren as formulated by William Byrd in his 2010 PhD thesis \cite{byrd2010relational}.

\subsection{GDL Translation to core.logic}
\subsubsection{Motivation}
Game Description Language is trivially derived from English. Below are the English and GDL descriptions of legality in Tic-Tac-Toe.
\begin{lstlisting}[frame=single, caption=The Legal relation for Tic-Tac-Toe expressed in English]
;It is Legal for player w to mark cell (m,n)
;if it is true that cell (m,n) is blank and
;it is true that it is player w's turn to
;move. If it is X's turn to move, O can only
;noop. If it is O's turn to move, X can only
;noop. 
\end{lstlisting}
\begin{lstlisting}[frame=single, caption=The Legal relation for Tic-Tac-Toe expressed in GDL]
(<= (legal ?w (mark ?m ?n))
    (true (cell ?m ?n b))
    (true (control ?w)))

(<= (legal X noop)
    (true (control O)))

(<= (legal O noop)
    (true (control X)))
\end{lstlisting}

Similarly, a corresponding core.logic relation can be derived from the GDL description.

\begin{lstlisting}[frame=single, caption=The Legal relation translated into core.logic]
(defn legal [role move]
 (conde
  [(fresh [?w ?m?n]
    (== role ?w)
    (== move [:mark ?m ?n])
    (true [:cell ?m ?n :b])
    (true [:control ?w]))]
  [(== role :X)
   (== move :noop)
   (true [:control :O])]
  [(== role :O)
   (== move :noop)
   (true [:control :X])]))
\end{lstlisting}

While the core.logic version is more verbose, each part has a direct correspondence(SP?) to part of the GDL description:
\begin{itemize}
\item The relation has an explicit declaration with argument names.
\item All the reflexive relations at the head of each GDL \textless= block are transformed into a series of unifications with the corresponding GDL arguments. 
\item All free variables in the GDL relation are now explicitly declared in a core.logic fresh block.
\item The three separate GDL relations are now combined in a single conde block. In the GDL description, there is an implicit disjunction between the relations. conde is a core.logic disjunction.
\end{itemize}

\subsubsection{Implementation}

\subsubsection{Results}
The translation algorithm was tested on four Game Description Language descriptions that can be found on the Stanford Gamemaster website: Tic-Tac-Toe, Chinkook, Chinese Checkers, and Pentago. These games are diverse in factors such as board structure, number of players, types of moves, synchronicity (??) of moves, etc (!!).
